---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## This document is by Aibek Mursalimov
# New project was created with an appropriate working directory
# Required packages were loaded

```{r, echo = TRUE}
library(dplyr)
library(ggplot2)
library(knitr)
```

## Loading and preprocessing the data

# For loading and reading the data

```{r, echo = TRUE}
knitr::opts_chunk$set(echo = TRUE)
unzip(zipfile = "activity.zip")
data_A <- read.csv("activity.csv")
str(data_A)
summary(data_A)
head(data_A)
```

# To further process data and remove missing values

```{r, echo = TRUE}
knitr::opts_chunk$set(echo = TRUE)
data_B <- na.omit(data_A)
```

## What is mean total number of steps taken per day?

# Creating the summary

```{r, echo = TRUE}
knitr::opts_chunk$set(echo = TRUE)
data_C <- group_by(data_B, date)
data_C <- summarize(data_C, steps=sum(steps))

summary(data_C)
```

# To draw a histogram

```{r, echo = TRUE}
knitr::opts_chunk$set(echo = TRUE)
qplot(steps, data=data_C)
```

# Calculating the mean and median

```{r, echo = TRUE}
knitr::opts_chunk$set(echo = TRUE)

mean(data_C$steps)

median(data_C$steps)
```

## What is the average daily activity pattern?

# creating a new data frame with the intervals

```{r, echo = TRUE}
knitr::opts_chunk$set(echo = TRUE)
data_D <- group_by(data_B, interval)
data_D <- summarize(data_D, steps=mean(steps))
```

# to create a histogram

```{r, echo = TRUE}
knitr::opts_chunk$set(echo = TRUE)

ggplot(data_D, aes(interval, steps)) + geom_line()
```

# Calculating the max

```{r, echo = TRUE}
knitr::opts_chunk$set(echo = TRUE)

data_D[data_D$steps==max(data_D$steps),]
```

## Imputing missing values

# processing data and filling missing values

```{r, echo = TRUE}
knitr::opts_chunk$set(echo = TRUE)
nrow(data_A)-nrow(data_B)

names(data_D)[2] <- "mean.steps"
act.impute <- merge(data_A, data_B)

act.impute$steps[is.na(act.impute$steps)] <- act.impute$mean.steps[is.na(act.impute$steps)]

act.day.imp <- group_by(act.impute, date)
act.day.imp <- summarize(act.day.imp, steps=sum(steps))
```

# creating a histogram

```{r, echo = TRUE}
knitr::opts_chunk$set(echo = TRUE)
qplot(steps, data=act.day.imp)
```

# Calculating mean and median

```{r, echo = TRUE}
knitr::opts_chunk$set(echo = TRUE)
mean(act.day.imp$steps)

median(act.day.imp$steps)
```


## Are there differences in activity patterns between weekdays and weekends?

# New data set
```{r, echo = TRUE}
knitr::opts_chunk$set(echo = TRUE)
act.impute$dayofweek <- weekdays(as.Date(act.impute$date))
act.impute$weekend <-as.factor(act.impute$dayofweek=="Saturday"|act.impute$dayofweek=="Sunday")
levels(act.impute$weekend) <- c("Weekday", "Weekend")

act.weekday <- act.impute[act.impute$weekend=="Weekday",]
act.weekend <- act.impute[act.impute$weekend=="Weekend",]

act.int.weekday <- group_by(act.weekday, interval)
act.int.weekday <- summarize(act.int.weekday, steps=mean(steps))
act.int.weekday$weekend <- "Weekday"
act.int.weekend <- group_by(act.weekend, interval)
act.int.weekend <- summarize(act.int.weekend, steps=mean(steps))
act.int.weekend$weekend <- "Weekend"

act.int <- rbind(act.int.weekday, act.int.weekend)
act.int$weekend <- as.factor(act.int$weekend)
```

# creating a plot

```{r, echo = TRUE}
knitr::opts_chunk$set(echo = TRUE)
ggplot(act.int, aes(interval, steps)) + geom_line() + facet_grid(weekend ~ .)
```

